Asteroseismology of GD358 with complex C/O core profiles

### 2016-02-18 msnew.tex
msnew.tex

%%\newcommand{\vdag}{(v)^\dagger}
\newcommand{\bvf}{Brunt-V\"ais\"al\"a }
\newcommand{\teff}{$\rm T_{eff}$}
\newcommand{\logg}{$\log$g }
\newcommand{\kic}{KIC }
\newcommand{\kicnsp}{KIC}
\newcommand{\sigrms}{$\sigma_{\rm RMS}$}
\newcommand{\mstar}{$M_*$}
\newcommand{\menv}{$M_{\rm env}$}
\newcommand{\mhe}{$M_{\rm He}$}
\newcommand{\xo}{$X_o$}
\newcommand{\qfm}{$q_{\rm fm}$}
\newcommand{\muHz}{\mbox{$\mu$Hz}}
\newcommand{\msun}{$M_\odot$}

@@

\begin{abstract}
We report on the analysis of XXXX numbers of hours of data collected on the pulsating helium atmosphere white dwarf star, GD358. From this data set added to archival data, we extract 15 frequencies, adding 4 modes to the 11 modes known previously. We present evidence that the 15 modes are $\ell=1$ modes, 13 of which belong to a consecutive sequence in $k$. We use models that include parameterized, complex carbon and oxygen core composition profiles to fit the periods.
\end{abstract}

@@

The subject of this paper, GD358 (V777 Her) is the brightest ($m_{\rm v}=13.7$) and best 
studied helium atmosphere white dwarf pulsator. It is located near the red edge of the 
instability strip, with \teff of 24000~K and \logg of 7.8 \citep{Koester2013}. GD358's 
pulsation spectrum contains a series of independent modes, many with complex 
frequency structure.  Models involving magnetic fields and oblique rotation have been 
proposed to explain that structure \citep{Montgomery10}.

Pulsating white dwarfs do not pulsate in all their available modes at all times.  
We have maintained an active observing program of this complex star in an effort 
to identify additional frequencies. These new observations have identified additional 
periods in GD358's pulsation spectrum, bringing the total known modes to 15. Thirteen of 
these modes belong to a consecutive $\ell=1$ sequence, the longest sequence observed 
in a DBV. This sequence is critical to constrain the asteroseismic fits and better define the
limitations of high precision white dwarf asteroseismology.

Paradoxally, among the DBVs that have enough detected periods to be fitted 
asteroseismically, GD358 is the only one that has not been analyzed using the complex 
C/O profiles adapted and parameterized from stellar evolution calculations
(e.g. {}\citet{Salaris97,Althaus05}). The most recent fits of GD358 \citep{Metcalfe03c} were 
performed using 11 observed modes and simple models where the oxygen abundance 
drops linearly from its central value to zero. We present a new detailed asteroseismic analysis of GD358, taking advantage of 
its long sequence of $\ell=1$ modes to better constrain the asteroseismic fits and 
define the limits of high precision white dwarf asteroseismology.

@@

The final reduction step is to combine the individual light curves (one example shown in Fig. \ref{tersk}) to produce 
complete light curves for GD358 for each observing season. We assume GD358 oscillates about 
a mean light level. This assumption allows us to carefully assess overlapping segments from 
different telescopes and identify and correct any vertical offsets. As discussed in detail 
in \citet{Provencal09}, we find no significant differences between the noise levels of 
amplitude spectra using: 1) the combination of all light curves including overlapping 
segments from different telescopes, 2) the combination of light curves where we retain 
only higher signal to noise observations in overlapping segments and 3) combining all light 
curves using data weighted by telescope aperture.

@@

Our goal is to compile a complete list of GD358's observed independent and combination frequencies 
to be used in a comprehensive asteroseismic analysis. GD358 is known for large scale changes in 
amplitudes and small but not insignificant frequency variations on a variety of timescales 
\citep{Provencal09}.  The amplitude and frequency variations evident in Fig.~\ref{fts} will 
produce artifacts in FTs, greatly complicating any frequency analysis.  
We reanalyzed all available archival data \citep{Provencal09, Kepler03, Winget94, Winget82} 
as well as our new observations to insure continuity in methodology.  We use 
{\sl Period04} \citep{Lenz05} for Fourier analysis and nonlinear least squares fitting to 
select statistically significant peaks in each FT.  As discussed in \citet{Provencal12}, 
we adopt the criterion that a peak must have an amplitude at least four times greater than the 
average noise level in the given frequency range. We define ``noise'' as the frequency-dependent 
average amplitude after prewhitening by the dominant frequencies. This is a conservative estimate, 
as it is impossible to ensure that all of the ``real'' frequencies are removed when determining
the noise level. This is certainly true for GD358, where the peaks above $\approx$ 2500 $\mu$Hz\ 
are combination frequencies \citep{Provencal09}.  Fig.~\ref{noise} displays the average noise 
as a function of frequency for the 2007-2015 observing seasons. A similar plot for the 
archival data is presented in Fig. 3 of \citet{Provencal09}. To confirm our uncertainty 
estimates, we calculated Monte Carlo simulations using the routine provided in {\sl Period04}.  
This routine generates a set of light curves using the original times, the fitted frequencies and 
amplitudes, and added Gaussian noise. A least squares fit is performed on each simulated light 
curve, with the distribution of fit parameters giving the uncertainties. Our Monte Carlo results 
are consistent with our average noise estimates.

@@

Fig.~\ref{schematic} presents a schematic representation of GD358's pulsation frequencies for all the 
available data. The features of asteroseismic importance are the localized bands between 900 and 2400
\muHz\ (1100 and 400 s). We interpret each band to represent a single radial overtone. The bands
themselves illustrate the second way in which GD358's frequencies vary. Each band consists of 
frequency detections from multiple observing seasons, but each band is significantly wider in 
frequency space than the frequency error of any single measurement, arguing that some process 
is acting on the frequencies. The simplest explanation is rotational splitting.  However, with the 
exception of the two highest frequency (shortest period) bands ($k=8$ and $k=9$ 
from \citet{Provencal09}), we find no clear evidence of multiplet structure in the frequency 
distributions of the lower frequency (longer period) bands. The lower frequency band widths are 
most likely the result of amplitude and/or frequency modulation that obscures any underlying 
signatures of rotational splitting.

@@

Fig. \ref{schematic} also shows that the widths of the bands is not constant, but changes from band to band.
\citet{Bell15} presents an interesting analysis of the hydrogen atmosphere pulsator (DAV) KIC4552982, 
in which they identify 17 bands of pulsation frequencies. This DAV follows the general pattern exhibited 
by GD358: the highest frequency (shortest period) mode shows evidence of rotational splitting, while the 
lower frequency (longer period) modes are complex bands. The authors note that this DAV's bands have 
different widths in frequency space, and that there may be astronomical significance to this. Although the 
observational timebases and coverage are quite different (20 continuous months for KIC4552982 vs 30 
incomplete years for GD358), our data presents the opportunity to investigate this for a DB pulsator.  
We measured the width of each of GD358's band. We define ''width'' as the difference between the 
lowest and highest frequency detected in each band. Fig. \ref{thick} presents the results. 
The widths of GD358's bands are commensurate with those found in KIC4552982.  We also find a general 
increase in width with decreasing frequency (increasing period), until we reach 
the mode at 1238 \muHz\ (807 s). This band has a width at least twice as wide as any other.  What 
is different about this mode???

\begin{figure}
 \epsscale{0.8}
 \plotone{modethick.eps}
 \caption{Width of each bank of power in Fig.~\ref{schematic}. We find an overall increase 
 in band thickness with decreasing frequency (increasing period).  The band at 807~s has a width
 at least twice as large as other bands.  
 \label{thick}}
\end{figure}

\subsection{Mode Identification}
Our current work has produced a well defined sequence of modes, adding to previous studies \citep{Montgomery10, Provencal09, Metcalfe00, Winget94}. The previous identification of these modes as 
a series of $l$=1 radial overtones is based mostly on the pulsation frequency distribution 
and limited spectroscopic analysis \citep{Kotak02, Castanheira05}. It is important to further 
investigate these identifications as we initiate an indepth asteroseismic investigation.

@@

The parameters for our best fit model based on the refined grid are listed in table \ref{fitt1} and the periods of that model in table \ref{gd}. The goodness of fit of the model is \sigrms$=4.22$~s. In Fig. \ref{ffit1}, we show the interior structure of the best fit model, with the corresponding \bvf frequency profile.

>>>>

%%\newcommand{\vdag}{(v)^\dagger}
\newcommand{\bvf}{Brunt-V\"ais\"al\"a }
\newcommand{\teff}{$\rm T_{eff}$}
\newcommand{\logg}{$\log$\ g }
\newcommand{\kic}{KIC }
\newcommand{\kicnsp}{KIC}
\newcommand{\sigrms}{$\sigma_{\rm RMS}$}
\newcommand{\mstar}{$M_*$}
\newcommand{\menv}{$M_{\rm env}$}
\newcommand{\mhe}{$M_{\rm He}$}
\newcommand{\xo}{$X_o$}
\newcommand{\qfm}{$q_{\rm fm}$}
\newcommand{\muHz}{\mbox{$\mu$Hz}}
\newcommand{\msun}{$M_\odot$}

@@

\begin{abstract}
We report on the analysis 33 years (1982-2015) of photometric observations of the pulsating
helium atmosphere white dwarf GD358.  The data set includes 1105.1 hours of new observations
from 2007-2015. From this data set, we extract 15 frequencies representing a sequence 
of radial pulsation modes, adding 4 modes to the 11 modes known previously. We present evidence 
that these 15 modes are $\ell=1$ modes, 13 of which belong to a consecutive sequence in radial overtone
$k$. We perform a detailed asteroseismic analysis using models that include parameterized, complex carbon and oxygen 
core composition profiles to fit the periods.Our results indicate.....

\end{abstract}

@@

The subject of this paper, GD358 (V777 Her) is the brightest ($m_{\rm v}=13.7$) and best 
studied helium atmosphere white dwarf pulsator. It is located near the red edge of the 
instability strip, with a spectroscopic temperature of \teff\ $=24000\pm500$ ~K and \logg\ $=7.8$ 
\citep{Koester2013}. GD358's pulsation spectrum contains a series of independent 
modes, many with complex frequency structure.  Models involving magnetic fields and oblique rotation 
have been proposed to explain that structure \citep{Montgomery10}.

Since the 2006 Whole Earth Telescope (WET) run reported in \citet{Provencal09}, 
we have maintained an active observing program of GD358 in an effort to identify additional
frequencies in this complex star. These new observations have identified additional 
periods in GD358's pulsation spectrum, bringing the total known modes to 15. Thirteen of 
these modes belong to a consecutive $\ell=1$ sequence, the longest sequence observed 
in a DBV.

Paradoxally, among the DBVs with enough detected periods to be fitted 
asteroseismically, GD358 is the only one that has not been analyzed using the complex 
C/O profiles adapted and parameterized from stellar evolution calculations
(e.g. {}\citet{Salaris97,Althaus05}). The most recent fits of GD358 \citep{Metcalfe03c} were 
performed using 11 observed modes and simple models where the oxygen abundance 
drops linearly from its central value to zero. We present a new detailed asteroseismic analysis of GD358, taking advantage of 
its long sequence of $\ell=1$ modes to better constrain the asteroseismic fits and 
define the limits of high precision white dwarf asteroseismology.

@@

The final reduction step is to combine the individual light curves (an example is shown in Fig. \ref{tersk}) to produce 
complete light curves for GD358 for each observing season. We assume GD358 oscillates about 
a mean light level. This assumption allows us to carefully assess overlapping segments from 
different telescopes and identify and correct any vertical offsets. As discussed in detail 
in \citet{Provencal09}, we find no significant differences between the noise levels of 
amplitude spectra using: 1) the combination of all light curves including overlapping 
segments from different telescopes, 2) the combination of light curves where we retain 
only higher signal to noise observations in overlapping segments and 3) combining all light 
curves using data weighted by telescope aperture.

@@

Our goal is to compile a complete list of GD358's observed independent and combination frequencies 
to be used in a comprehensive asteroseismic analysis. GD358 is known for large scale changes in 
amplitudes and small but not insignificant frequency variations on a variety of timescales 
\citep{Provencal09}.  The amplitude and frequency variations evident in Fig.~\ref{fts} will 
produce artifacts in FTs, greatly complicating any frequency analysis.  
We reanalyzed all available archival data \citep{Provencal09, Kepler03, Winget94, Winget82} 
as well as our new observations to insure continuity in methodology.  We use 
{\sl Period04} \citep{Lenz05} for Fourier analysis and nonlinear least squares fitting to 
select statistically significant peaks in each FT.  As discussed in \citet{Provencal12}, 
we adopt the criterion that a peak must have an amplitude at least four times greater than the 
average noise level in the given frequency range. We define ``noise'' as the frequency-dependent 
average amplitude after prewhitening by the dominant frequencies. This is a conservative estimate, 
as it is impossible to ensure that all of the ``real'' frequencies are removed when determining
the noise level. This is certainly true for GD358, where the peaks above $\approx$ 2500 $\mu$Hz\ 
are combination frequencies (see Section \ref{mode}, \citet{Provencal09}).  Fig.~\ref{noise} displays the average noise 
as a function of frequency for the 2007-2015 observing seasons. A similar plot for the 
archival data is presented in Fig. 3 of \citet{Provencal09}. To confirm our uncertainty 
estimates, we calculated Monte Carlo simulations using the routine provided in {\sl Period04}.  
This routine generates a set of light curves using the original times, the fitted frequencies and 
amplitudes, and added Gaussian noise. A least squares fit is performed on each simulated light 
curve, with the distribution of fit parameters giving the uncertainties. Our Monte Carlo results 
are consistent with our average noise estimates.

@@

Fig.~\ref{schematic} presents a schematic representation of GD358's 
independent frequencies detected between 1982 and 2015.  
The features of asteroseismic importance are the localized bands between 900 and 2400
\muHz\ (1100 and 400 s). We interpret each band to represent a single radial overtone. The bands
themselves illustrate the second way in which GD358's frequencies vary. Each band consists of 
frequency detections from multiple observing seasons, but each band is significantly wider in 
frequency space than the frequency error of any single measurement, arguing that some process 
is acting on the frequencies. The simplest explanation is rotational splitting.  However, with the 
exception of the two highest frequency (shortest period) bands ($k=8$ and $k=9$ 
from \citet{Provencal09}), we find no clear evidence of multiplet structure in the frequency 
distributions of the lower frequency (longer period) bands (Fig.~\ref{rot}). The lower frequency band widths are 
most likely the result of amplitude and/or frequency modulation that obscures any underlying 
signatures of rotational splitting.

@@

Fig. \ref{schematic} also shows that the widths of the bands is not constant, but changes from band to band.
\citet{Bell15} presents an interesting analysis of the hydrogen atmosphere pulsator (DAV) KIC4552982, 
in which they identify 17 bands of pulsation frequencies. This DAV follows the general pattern exhibited 
by GD358: the highest frequency (shortest period) mode shows evidence of rotational splitting, while the 
lower frequency (longer period) modes are complex bands. The authors note that this DAV's bands have 
different widths in frequency space, and that there may be astronomical significance to this. Although the 
observational timebases and coverage are quite different (20 continuous months for KIC4552982 vs 30 
incomplete years for GD358), our data presents the opportunity to investigate this for a DB pulsator.  
We measured the width of each of GD358's band. We define ''width'' as the difference between the 
lowest and highest frequency detected in each band. Fig. \ref{thick} presents the results. 
The widths of GD358's bands are commensurate with those found in KIC4552982.  We also find a general 
increase in width with decreasing frequency (increasing period), until we reach 
the mode at 1238 \muHz\ (807 s). This band has a width at least twice as wide as any other.

A second interesting pattern is given by the distribution of amplitudes for the modes.  In particular, the 
two modes at 2007.6 and 1857.7 \muHz\ have never been observed at amplitudes above 4 mma.  Interestingly, the 
mode at 1741.5 \muHz\ was not observed at large amplitude prior to 2006.  This ``zone of avoidance'' is telling 
us something about the structure of the star. Lower frequency modes preferentially sample the outer layers of 
the star.  
\begin{figure}
 \epsscale{0.8}
 \plotone{modethick.eps}
 \caption{Width of each bank of power in Fig.~\ref{schematic}. We find an overall increase 
 in band thickness with decreasing frequency (increasing period).  The band at 807~s has a width
 at least twice as large as other bands.  
 \label{thick}}
\end{figure}

\subsection{Mode Identification}\label{mode}
Our current work has produced a well defined sequence of modes, adding to previous studies \citep{Montgomery10, Provencal09, Metcalfe00, Winget94}. The previous identification of these modes as 
a series of $l$=1 radial overtones is based mostly on the pulsation frequency distribution 
and limited spectroscopic analysis \citep{Kotak02, Castanheira05}. It is important to further 
investigate these identifications as we initiate an indepth asteroseismic investigation.

@@

The parameters for our best fit model based on the refined grid are listed in table \ref{fitt1} and the periods of that model in 
Table \ref{gd}. The goodness of fit of the model is \sigrms$=4.22$~s. In Fig. \ref{ffit1}, we show the interior structure of 
the best fit model, with the corresponding \bvf frequency profile.
